---
title: "Matrix Algebra"
output: html_notebook
---
# Motivating Examples 

## Falling objects
```{r}
set.seed(1)
g <- 9.8
n <- 25
tt <- seq(0,3.4,len=n)
d <- 56.67 - 0.5*g*tt*2 + rnorm(n,sd=1)
plot(tt,d,ylab="Distance in meters",xlab="Time in seconds")
```
By looking at the plot one can deduce that the position should follow a parabola. We can model by - 

$$ Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_{i}^{2} + \epsilon, i = 1......n $$

## Father and son heights 
```{r}
library(UsingR)
x = father.son$fheight
y = father.son$sheight
plot(x,y,xlab="Father's height", ylab="Son's height")
```
The son's height do seem to increase linearly with father's heights. In this case we chose a linear model - 

$$ Y_i = \beta_0 + \beta_1 x_i + \epsilon, i = 1,......,N $$
## Random samples from multiple populations 
We are interested to determine if the diet has an effect on weight. 
```{r}
dat <- read.csv("femaleMiceWeights.csv")
stripchart(Bodyweight~Diet,data=dat,vertical=TRUE,method="jitter",pch=1,main="Mice weights")
```

```{r}
g <- 9.8 
n <- 25 
tt <- seq(0,3.4,len=n)
f <- 56.67 - 0.5*g*tt^2
y <- f + rnorm(n,sd=1)
plot(tt,y,ylab = "Distance in meters", xlab = "Time in seconds")
lines(tt,f,col=2)
```

## The lm function 
```{r}
tt2 <- tt^2
fit <- lm(y~tt+tt2)
summary(fit)$coef
```
## The least squares estimate (LSE)
```{r}
rss <- function(Beta0,Beta1,Beta2){
  r <- y - (Beta0+Beta1*tt+Beta2*tt^2)
  return(sum(r^2))
}
```

For any three dimensional vector we get an RSS. Here is a plot as a function of $$\beta_2$$ when other two are fixed. 

```{r}
Beta2s <- seq(-10,0,len=100)
plot(Beta2s,sapply(Beta2s,rss,Beta0=55,Beta1=0),
     ylab="RSS",xlab="Beta2",type="l")
## another curve fixing another pair
Beta2s <- seq(-10,0,len=100)
lines(Beta2s,sapply(Beta2s,rss,Beta0=65,Beta1=0),col=2)
```
## Exercises 
```{r}
library(UsingR)
data("father.son",package="UsingR")
```
```{r}
# Question 1 
mean(father.son$sheight)
```
```{r}
# Question 2 

```
## Vectors, Matrices and Scalars 
```{r}
library(UsingR)
y = father.son$fheight
head(y)
```
```{r}
n <- 25 
tt <- seq(0,3.4,len=n)
X <- cbind(X1=tt,X2=tt^2)
head(X)
```
```{r}
N <-100 
p <- 5
X <- matrix(1:(N*p),N,p)
head(X)
```
```{r}
N <- 100; p<- 5
X <- matrix(1:(N*p),N,p,byrow=TRUE)
head(X)
```
## Exercises 
```{r}
"1."
a <- c(1,5,2,4)
b <- rnorm(10)
X = matrix(1:1000, 100,10)
X[25,4]

```
```{r}
"2."
x <- 1:10
cbind(x,2*x,3*x,4*x,5*x)
```