---
title: "Matrix Algebra"
output: html_notebook
---
# Motivating Examples 

## Falling objects
```{r}
set.seed(1)
g <- 9.8
n <- 25
tt <- seq(0,3.4,len=n)
d <- 56.67 - 0.5*g*tt*2 + rnorm(n,sd=1)
plot(tt,d,ylab="Distance in meters",xlab="Time in seconds")
```
By looking at the plot one can deduce that the position should follow a parabola. We can model by - 

$$ Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_{i}^{2} + \epsilon, i = 1......n $$

## Father and son heights 
```{r}
library(UsingR)
x = father.son$fheight
y = father.son$sheight
plot(x,y,xlab="Father's height", ylab="Son's height")
```
The son's height do seem to increase linearly with father's heights. In this case we chose a linear model - 

$$ Y_i = \beta_0 + \beta_1 x_i + \epsilon, i = 1,......,N $$
## Random samples from multiple populations 
We are interested to determine if the diet has an effect on weight. 
```{r}
dat <- read.csv("femaleMiceWeights.csv")
mypar(1,1)
stripchart(Bodyweight~Diet,data=dat,vertical=TRUE,method="jitter",pch=1,main="Mice weights")
```

```{r}
g <- 9.8 
n <- 25 
tt <- seq(0,3.4,len=n)
f <- 56.67 - 0.5*g*tt^2
y <- f + rnorm(n,sd=1)
plot(tt,y,ylab = "Distance in meters", xlab = "Time in seconds")
lines(tt,f,col=2)
```

## The lm function 